---
output:
  word_document: default
  html_document: default
---

```{r warning = FALSE}

 #excel with Maturity from 31.1.2018 till 01.02.2016

library(readxl)
yields  <- read_excel("C:/Users/User/Desktop/data.xlsx")

```


```{r warning = FALSE}
matplot(yields, type='l', ylim = c(0,12),ylab = 'rate', xlab = 'Time') #yield curve

```



```{r warning = FALSE}
returns = data.frame(diff(as.matrix(yields)))*(-1) #daily zero-coupons increments
head(returns)

```

Сorrelation

```{r warning = FALSE}
options(digits = 2)
cor(returns) 

```

Сovariance

```{r warning = FALSE}
options(digits = 2)
cov(returns)

```



```{r warning = FALSE}
#necessary to standardize the variables because PCA is sensitive to the scale of the data
model = prcomp(returns, scale = TRUE, center = TRUE) 
```

Data frame which provides the data for the principal components analysis

```{r warning = FALSE}
x = data.frame(model$x)
head(x)
tail(x)
```


The matrix of variable loadings (i.e., a matrix whose columns contain the eigenvectors)

```{r warning = FALSE}

loadings = data.frame(model$rotation) 
head(loadings)
tail(loadings)
```


Standard deviation - this is simply the eigenvalues in our case since the data has been centered and scaled (standardized)

Proportion of Variance - the amount of variance the component accounts for in the data

Cumulative Proportion - the accumulated amount of explained variance

Three principal components can explain almost 97% percent of the variance. Eigenvalues of first two components are larger than 1. Eigenvalue of the third component is almost 1. 

```{r}
summary(model)
```


```{r warning = FALSE}
par(mfrow = c(1,2))
barplot(model$sdev^2, main = 'Eigenvalues of each component')
barplot(cumsum(model$sdev^2)/sum(model$sdev^2), main = 'Cumulative Explained Variance', ylab = 'Variance Explained')
```

```{r warning = FALSE}
plot(loadings$PC1, type = "line") #convexity
plot(loadings$PC2, type = "line") #slope change
plot(loadings$PC3, type = "line") #level change

```




